manuscript_figures

### 2024-06-25 scripts/manuscript_figures.Rmd
updated and added temp, rate of change, kelp biomass plots

# Make figure 3 and run statistical analysis
```{r}
# Aggregate the data by year, site, and transect and compute mean values
aggregated.data.morph <- merged.dat.morph %>%
  group_by(YEAR, SITE, TRANSECT) %>%
  summarize(mean_fish_biomass = mean(Fish.Biomass..g.m2., na.rm = TRUE),
            mean_kelp_biomass = mean(average_kelp_biomass, na.rm = TRUE)) %>%
  ungroup()

@@

  labs(x = "Log Kelp Biomass (g/m2)", y = "Log Fish biomass (g/m2)",
       title = "Fish Biomass vs Kelp Biomass") +
  theme_classic() +
  theme(plot.margin = margin(t = 10, r = 30, b = 10, l = 10, unit = "pt")) +
  coord_cartesian(clip = 'off');figure.3

@@

# Figure 4: organize rugosity data and make linear regression figure for fish biomass and rugosity
```{r}
# Calculate mean rugosity complexity scores across depths for each transect, grouped by site and year
mean_rugosity <- rugosity %>%
  group_by(SITE, YEAR, TRANSECT) %>%
  summarize(mean_rugosity = mean(c(X0M, X4M, X8M, X12M, X16M, X20M, X24M, X28M, X32M, X36M, X40M), na.rm = TRUE))

# edited control transcet name
mean_rugosity$TRANSECT[mean_rugosity$TRANSECT == "Control 2"] <- "2"

@@

# Run all above models without rugosity

```{r}

@@

@@

>>>>

# Make figure 3 fish biomass x kelp biomass and run statistical analysis
```{r}
# Aggregate the data by year, site, and transect and compute mean values
aggregated.data.morph <- merged.dat.morph %>%
  group_by(YEAR, SITE, TRANSECT) %>%
  summarize(mean_fish_biomass = mean(Fish.Biomass..g.m2., na.rm = TRUE),
            mean_kelp_biomass = mean(average_kelp_biomass, na.rm = TRUE)) %>%
  ungroup()

@@

  labs(x = expression("Log Kelp Biomass (g/" ~ m^2 ~ ")"), y = expression("Log Fish Biomass (g/" ~ m^2 ~ ")")) +
  theme_classic() +
  theme(plot.margin = margin(t = 10, r = 30, b = 10, l = 10, unit = "pt")) +
  coord_cartesian(clip = 'off') +
  theme(
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.x = element_text(size = 18),
    axis.text.y = element_text(size = 18));figure.3

@@

# Make kelp biomass time series (a) 
```{r}
# Function to calculate standard error
std_error <- function(x) {
  sd(x, na.rm = TRUE) / sqrt(length(na.omit(x)))
}

# Aggregate the data by year and site and compute mean values and standard error
aggregated.kelp.morph <- morph.filt %>%
  group_by(YEAR, SITE) %>%
  summarize(
    mean_kelp_biomass = mean(average_kelp_biomass, na.rm = TRUE),
    se_kelp_biomass = std_error(average_kelp_biomass)
  ) %>%
  ungroup()

# Transform mean kelp biomass and standard error to log scale
aggregated.kelp.morph <- aggregated.kelp.morph %>%
  mutate(
    log_mean_kelp_biomass = log(mean_kelp_biomass + 1.1),
    log_se_kelp_biomass = log(mean_kelp_biomass + 1.1 + se_kelp_biomass) - log(mean_kelp_biomass + 1.1)
  )

# Define shapes for sites
shape_values <- c("Fort Wetherill" = 16, "King's Beach" = 17) # Circle for Fort Wetherill, Triangle for King's Beach

# Time series plot for kelp biomass with standard error bars from 2016-2023
kelp.timeseries.2016 <- ggplot(aggregated.kelp.morph, aes(x = YEAR, y = log_mean_kelp_biomass, shape = SITE)) +
  geom_line(aes(group = SITE), color = "black") +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = log_mean_kelp_biomass - log_se_kelp_biomass, 
                    ymax = log_mean_kelp_biomass + log_se_kelp_biomass), 
                width = 0.2) +
  scale_shape_manual(values = shape_values) +
  labs(x = "Year", y = expression("Log Mean Kelp Biomass (g/" ~ m^2 ~ ")"), shape = "Site") +
  theme_bw() +
  scale_x_continuous(breaks = unique(aggregated.kelp.morph$YEAR)) +
  theme(
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.x = element_text(size = 18),
    axis.text.y = element_text(size = 18),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18)
  );kelp.timeseries.2016

# Save the plot to a file
ggsave("output/figures/kelp_timeseries_2016.png", plot = kelp.timeseries.2016, width = 8, height = 6)

# Filter for the years 2019-2023
filtered_data <- aggregated.kelp.morph %>%
  filter(YEAR >= 2019 & YEAR <= 2023)

# Time series plot for kelp biomass with standard error bars for 2019-2023
kelp.timeseries.2019_2023 <- ggplot(filtered_data, aes(x = YEAR, y = log_mean_kelp_biomass, shape = SITE)) +
  geom_line(aes(group = SITE), color = "black") +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = log_mean_kelp_biomass - log_se_kelp_biomass, 
                    ymax = log_mean_kelp_biomass + log_se_kelp_biomass), 
                width = 0.2) +
  scale_shape_manual(values = shape_values) +
  labs(x = "Year", y = expression("Log Mean Kelp Biomass (g/" ~ m^2 ~ ")"), shape = "Site") +
  theme_bw() +
  theme(
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.x = element_text(size = 18),
    axis.text.y = element_text(size = 18),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18)
  );kelp.timeseries.2019_2023

# Save the plot to a file
ggsave("output/figures/kelp_timeseries_2019.png", plot = kelp.timeseries.2019_2023, width = 8, height = 6)

# Rate of change (b) by site figure
```{r}
# Fit the linear model
lm_model <- lm(average_kelp_biomass ~ YEAR, data = morph.filt)
#check normality
qqPlot(residuals(lm_model)) #data is normal no need to transform

# Fit log transformed linear model
lm_model <- lm(log(average_kelp_biomass + 1.1) ~ YEAR, data = morph.filt)
#check normality
qqPlot(residuals(lm_model)) #data is normal no need to transform

# Convert YEAR to numeric (if necessary)
morph.filt$YEAR <- as.numeric(as.character(morph.filt$YEAR))

# Define shapes for sites
shape_values <- c("Fort Wetherill" = 16, "King's Beach" = 17)  # Circle for Fort Wetherill, Triangle for King's Beach

# Fit linear regression models for each SITE separately
slope_results <- morph.filt %>%
  mutate(log_avg_kelp_biomass = log(average_kelp_biomass + 1.1)) %>%
  group_by(SITE) %>%
  summarize(
    mean_slope = coef(lm(log_avg_kelp_biomass ~ YEAR))[2],  # Mean slope
    se_slope = summary(lm(log_avg_kelp_biomass ~ YEAR))$coefficients[2, "Std. Error"],  # Standard error of slope
    mean_log_avg_kelp_biomass = mean(log_avg_kelp_biomass)  # Mean log average kelp biomass
  )

# Calculate rate of change per year
slope_results <- slope_results %>%
  mutate(
    rate_of_change_per_year = mean_slope / mean_log_avg_kelp_biomass,  # Rate of change per year
    se_rate_of_change = se_slope / mean_log_avg_kelp_biomass  # Standard error of rate of change per year
  )

# Plotting the mean rates of change with error bars (standard error horizontally)
rate_of_change <- ggplot(slope_results, aes(x = rate_of_change_per_year, y = mean_log_avg_kelp_biomass, shape = SITE)) +
  geom_point(size = 3) +
  geom_errorbarh(aes(xmin = rate_of_change_per_year - se_rate_of_change, xmax = rate_of_change_per_year + se_rate_of_change), height = 0.2) +
  labs(
    x = expression("Rate of Change ("~yr^-1 ~")"),
    y = expression("Log Mean Kelp Biomass (g/"~m^2~")"), shape = "Site") +
  theme_bw() +
  scale_shape_manual(values = shape_values, guide = guide_legend(title = "Site")) +   # Use shape values and customize legend
  theme(
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.x = element_text(size = 18),
    axis.text.y = element_text(size = 18),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18));rate_of_change

# Plot two panels together using patchwork
```{r}
# Combine the plots with patchwork for all years of kelp and rate of change
kelp.rate.2016 <- kelp.timeseries.2016 | rate_of_change +
  plot_annotation(tag_levels = 'A') + 
  theme(plot.tag = element_text(size = 20, face = "bold"));kelp.rate.2016

# Save the combined plot
ggsave("output/figures/kelp.2016.rate.png", kelp.rate.2016, width = 20, height = 8, units = "in", dpi = 300)

# Combine the plots with patchwork for all years of kelp and rate of change
kelp.rate.2019 <- kelp.timeseries.2019_2023 | rate_of_change +
  plot_annotation(tag_levels = 'A') + 
  theme(plot.tag = element_text(size = 20, face = "bold"));kelp.rate.2019

# Save the combined plot
ggsave("output/figures/kelp.2019.rate.png", kelp.rate.2019, width = 20, height = 8, units = "in", dpi = 300)

# Organize rugosity data and make linear regression figure for fish biomass and rugosity
```{r}
# Calculate mean rugosity complexity scores across depths for each transect, grouped by site and year
mean_rugosity <- rugosity %>%
  group_by(SITE, YEAR, TRANSECT) %>%
  summarize(mean_rugosity = mean(c(X0M, X4M, X8M, X12M, X16M, X20M, X24M, X28M, X32M, X36M, X40M), na.rm = TRUE))

@@

# Run all above models without rugosity
```{r}

@@

# Make line plot for temperature time series by site per month
```{r}
# Calculate average temperature per month and site
avg_temp <- temperature_data %>%
  mutate(MonthAbbrev = month.abb[Month]) %>%  # Add month abbreviations
  group_by(MonthAbbrev, Site) %>%
  summarize(avg_temp = mean(Temp))

# Map site codes to full names
avg_temp$Site <- ifelse(avg_temp$Site == "FW", "Fort Wetherill", "King's Beach")

# Plotting average temperature by site as lines and points per month
temperature_plot <- ggplot(avg_temp, aes(x = MonthAbbrev, y = avg_temp, color = Site, group = Site)) +
  geom_line(size = 1.5) +  # Lines for average temperature by site
  geom_point(size = 3) +  # Points for mean values by site
  labs(x = "Month", y = expression("Average Temperature (°C)"), color = "Site") +
  scale_color_manual(values = c("Fort Wetherill" = "grey", "King's Beach" = "black")) +  # Custom colors
  theme_bw()+
    theme(
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.x = element_text(size = 18),
    axis.text.y = element_text(size = 18),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18));temperature_plot

# Save the plot to a file
ggsave("output/figures/temp.plot.png", plot = temperature_plot, width = 20, height = 10)
```

@@

Production to biomass ratios (P:B) for each fish species
```{r}
